---
title: "How Bike-share ride Members differ from Casual Bike-share Subscribers"

date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## STEP 1: UPLOADING THE DATA

The data was uploaded using and the data structure was specified. For example, column types were set to datetime(dttm), double (dbl), strings, etc.


```{r STEP1, message=FALSE, warning=FALSE}
library(tidyverse)
Q1_2020<- read_csv("cyclist/2020 DATA/Divvy_Trips_2020_Q1.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202004<- read_csv("cyclist/2020 DATA/202004-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202005<- read_csv("cyclist/2020 DATA/202005-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202006<- read_csv("cyclist/2020 DATA/202006-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202007<- read_csv("cyclist/2020 DATA/202007-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202008<- read_csv("cyclist/2020 DATA/202008-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202009<- read_csv("cyclist/2020 DATA/202009-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202010<- read_csv("cyclist/2020 DATA/202010-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202011<- read_csv("cyclist/2020 DATA/202011-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202012<- read_csv("cyclist/2020 DATA/202012-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202101<- read_csv("cyclist/2021 DATA/202101-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202102<- read_csv("cyclist/2021 DATA/202101-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202103<- read_csv("cyclist/2021 DATA/202103-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202104<- read_csv("cyclist/2021 DATA/202104-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202105<- read_csv("cyclist/2021 DATA/202105-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202106<- read_csv("cyclist/2021 DATA/202106-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202107<- read_csv("cyclist/2021 DATA/202108-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202108<- read_csv("cyclist/2021 DATA/202108-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202109<- read_csv("cyclist/2021 DATA/202109-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202110<- read_csv("cyclist/2021 DATA/202110-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202111<- read_csv("cyclist/2021 DATA/202111-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))
X202112<- read_csv("cyclist/2021 DATA/202112-divvy-tripdata.csv", 
                   col_types = cols(ride_id = col_character(), 
                                    rideable_type = col_character(), 
                                    started_at = col_datetime(format = "%d/%m/%Y %H:%M"), ended_at = col_datetime(format = "%d/%m/%Y %H:%M"), start_station_name = col_character(), start_station_id = col_double(), end_station_name = col_character(), end_station_id = col_double(), start_lat = col_double(), start_lng = col_double(), end_lat = col_double(), end_lng = col_double(), member_casual = col_character(), ride_length = col_time(format = "%H:%M:%S"), weekday = col_double()))

q2_2019 <- read_csv("Divvy_Trips_2019_Q2.csv")
q3_2019 <- read_csv("Divvy_Trips_2019_Q3.csv")
q4_2019 <- read_csv("Divvy_Trips_2019_Q4.csv")
q1_2019 <- read_csv("Divvy_Trips_2019_Q1.csv")

## Check the structures to ensure coltypes are in order.

str(q1_2019)
str(q2_2019)
str(q3_2019)
str(q4_2019)
```

## Cleaning the Data

The data in 2019 was renamed.

```{r Data Cleaning, message=FALSE, warning=FALSE}
(Q4_2019 <- rename(q4_2019
                   ,ride_id = trip_id
                   ,rideable_type = bikeid
                   ,started_at = start_time
                   ,ended_at = end_time
                   ,start_station_name = from_station_name
                   ,start_station_id = from_station_id
                   ,end_station_name = to_station_name
                   ,end_station_id = to_station_id
                   ,member_casual = usertype))
(Q3_2019 <- rename(q3_2019
                   ,ride_id = trip_id
                   ,rideable_type = bikeid
                   ,started_at = start_time
                   ,ended_at = end_time
                   ,start_station_name = from_station_name
                   ,start_station_id = from_station_id
                   ,end_station_name = to_station_name
                   ,end_station_id = to_station_id
                   ,member_casual = usertype))
(Q2_2019 <- rename(q2_2019
                   ,ride_id = "01 - Rental Details Rental ID"
                   ,rideable_type = "01 - Rental Details Bike ID"
                   ,started_at = "01 - Rental Details Local Start Time"
                   ,ended_at = "01 - Rental Details Local End Time"
                   ,start_station_name = "03 - Rental Start Station Name"
                   ,start_station_id = "03 - Rental Start Station ID"
                   ,end_station_name = "02 - Rental End Station Name"
                   ,end_station_id = "02 - Rental End Station ID"
                   ,member_casual = "User Type"))
(Q1_2019 <- rename(q1_2019
                   ,ride_id = "trip_id"
                   ,rideable_type = "bikeid"
                   ,started_at = "start_time"
                   ,ended_at = "end_time"
                   ,start_station_name = "from_station_name"
                   ,start_station_id = "from_station_id"
                   ,end_station_name = "to_station_name"
                   ,end_station_id = "to_station_id"
                   ,member_casual = "usertype"))

```

## DATA MERGING
 Data of all the month in 2020 and 2021 will be merged into one as all_trips_2020 and all_trips_2021 respectively and all the quarters in 2019 as all_trip_2019.
 
```{r DATA MERGING, message=FALSE, warning=FALSE}
all_trips_2019 <- bind_rows(Q1_2019,Q2_2019,Q3_2019,Q4_2019)
all_trips_2020 <- bind_rows(Q1_2020, X202004, X202005, X202006, 
                            X202007, X202008, X202009, X202010,
                            X202011, X202012)
all_trips_2021 <- bind_rows(X202101, X202102, 
                       X202103, X202104, X202105, X202106,
                       X202107, X202108, X202109,
                       X202110, X202111, X202112)
```

## DATA CLEANING 2
The merged data, all_trips_2019,all_trips_2020,and all_trips_2021 will be cleaned using these code chunk:

```{r DATA CLEANING 2, message=FALSE, warning=FALSE}
library(lubridate)
## change the string in rideable_type of 2019 to suit rideable_type_2020 and 2021)
mutate(all_trips_2019, ride_id = as.character(ride_id, rideable_type = as.character(rideable_type)))
all_trips_2019 <- all_trips_2019 %>%
  mutate(member_casual = recode(member_casual
                                ,"Subscriber" = "member"
                                ,"Customer" = "casual"))

## CREATE NEW COLS SUCH AS DATE,DAY, MONTH, YEAR AND DAY OF WEEK
all_trips_2019$date <- as.Date(all_trips_2019$started_at) #The default format is yyyy-mm-dd
all_trips_2019$month <- format(as.Date(all_trips_2019$date), "%m")
all_trips_2019$day <- format(as.Date(all_trips_2019$date), "%d")
all_trips_2019$year <- format(as.Date(all_trips_2019$date), "%Y")
all_trips_2019$day_of_week <- format(as.Date(all_trips_2019$date), "%A")


all_trips_2019$ride_length <- difftime(all_trips_2019$ended_at,all_trips_2019$started_at)
is.factor(all_trips_2019$ride_length)
all_trips_2019$ride_length <- as.numeric(as.character(all_trips_2019$ride_length))
is.numeric(all_trips_2019$ride_length)
all_trips_2020$ride_length <- difftime(all_trips_2020$ended_at,all_trips_2020$started_at)
is.factor(all_trips_2020$ride_length)
all_trips_2020$ride_length <- as.numeric(as.character(all_trips_2020$ride_length))
is.numeric(all_trips_2020$ride_length)
all_trips_2021$ride_length <- difftime(all_trips_2021$ended_at,all_trips_2021$started_at)
is.factor(all_trips_2021$ride_length)
all_trips_2021$ride_length <- as.numeric(as.character(all_trips_2021$ride_length))
is.numeric(all_trips_2021$ride_length)

## remove bad data
all_trips_2019_v2 <- all_trips_2019[!(all_trips_2019$start_station_name == "HQ QR" | all_trips_2019$ride_length<0),]
all_trips_2020_v2 <- all_trips_2020[!(all_trips_2020$start_station_name == "HQ QR" | all_trips_2020$ride_length<0),]
all_trips_2021_v2 <- all_trips_2021[!(all_trips_2021$start_station_name == "HQ QR" | all_trips_2021$ride_length<0),]
## CREATE NEW COLS SUCH AS DATE,DAY, MONTH, YEAR AND DAY OF WEEK
all_trips_2020_v2$date <- as.Date(all_trips_2020_v2$started_at) 
#The default format is yyyy-mm-dd
all_trips_2020_v2$month <- format(as.Date(all_trips_2020_v2$date), "%m")
all_trips_2020_v2$day <- format(as.Date(all_trips_2020_v2$date), "%d")
all_trips_2020_v2$year <- format(as.Date(all_trips_2020_v2$date), "%Y")
all_trips_2020_v2$day_of_week <- format(as.Date(all_trips_2020_v2$date), "%A")

all_trips_2021_v2$date <- as.Date(all_trips_2021_v2$started_at) 
#The default format is yyyy-mm-dd
all_trips_2021_v2$month <- format(as.Date(all_trips_2021_v2$date), "%m")
all_trips_2021_v2$day <- format(as.Date(all_trips_2021_v2$date), "%d")
all_trips_2021_v2$year <- format(as.Date(all_trips_2021_v2$date), "%Y")
all_trips_2021_v2$day_of_week <- format(as.Date(all_trips_2021_v2$date), "%A")

## select only the needed by deselecting unwanted data
all_trips_2019_v3 <- all_trips_2019_v2 %>%
  select(-c("birthyear", "gender", "01 - Rental Details Duration In Seconds Uncapped", "05 - Member Details Member Birthday Year", "Member Gender", "tripduration"))

all_trips_2020_v3 <- all_trips_2020_v2 %>%
  select(-c("start_lat", "start_lng", "end_lat", "end_lng", "weekday", "Weekday"))

all_trips_2021_v3 <- all_trips_2021_v2 %>%
  select(-c("start_lat", "start_lng", "end_lat", "end_lng", "weekday"))

```

## DATA ANALYSIS
This will be done separately for each of the year

```{r ANALYSIS OF THE CLEANED DATA, message=FALSE, warning=FALSE}

#YEAR 2019
summary(all_trips_2019_v3$ride_length)
# Compare members and casual users
aggregate(all_trips_2019_v3$ride_length ~ all_trips_2019_v3$member_casual, FUN = mean)
aggregate(all_trips_2019_v3$ride_length ~ all_trips_2019_v3$member_casual, FUN = median)
aggregate(all_trips_2019_v3$ride_length ~ all_trips_2019_v3$member_casual, FUN = max)
aggregate(all_trips_2019_v3$ride_length ~ all_trips_2019_v3$member_casual, FUN = min)
# The average ride time by each day for members vs casual users
aggregate(all_trips_2019_v3$ride_length ~ all_trips_2019_v3$member_casual + all_trips_2019_v3$day_of_week, FUN = mean)
# Notice that the days of the week are out of order. I'll fix that.
all_trips_2019_v3$day_of_week <- ordered(all_trips_2019_v3$day_of_week, levels=c("Sunday", "Monday",
"Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
# The average ride time by each day for members vs casual users
aggregate(all_trips_2019_v3$ride_length ~ all_trips_2019_v3$member_casual + all_trips_2019_v3$day_of_week, FUN = mean)
# analyze ridership data by type and day_of_week
all_trips_2019_v3 %>%
  group_by(member_casual, day_of_week) %>% #groups by user type and day_of_week
  summarise(number_of_rides = n() #calculates the number of rides and average duration
            ,average_duration = mean(ride_length)) %>% # calculates the average duration
  arrange(member_casual, day_of_week) # sorts
# visualize the number of rides by rider type
all_trips_2019_v3 %>%
  group_by(member_casual, day_of_week) %>%
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>%
  arrange(member_casual, day_of_week, na = TRUE) %>%
  ggplot(aes(x = day_of_week, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")
# Create a visualization for average duration
all_trips_2019_v3 %>%
  group_by(member_casual, day_of_week) %>%
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>%
  arrange(member_casual, day_of_week) %>%
  ggplot(aes(x = day_of_week, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge")
counts_2019 <- aggregate(all_trips_2019_v3$ride_length ~ all_trips_2019_v3$member_casual + all_trips_2019_v3$day_of_week, FUN = mean)


#YEAR 2020
summary(all_trips_2020_v3$ride_length)
# Compare members and casual users
aggregate(all_trips_2020_v3$ride_length ~ all_trips_2020_v3$member_casual, FUN = mean)
aggregate(all_trips_2020_v3$ride_length ~ all_trips_2020_v3$member_casual, FUN = median)
aggregate(all_trips_2020_v3$ride_length ~ all_trips_2020_v3$member_casual, FUN = max)
aggregate(all_trips_2020_v3$ride_length ~ all_trips_2020_v3$member_casual, FUN = min)
# The average ride time by each day for members vs casual users
aggregate(all_trips_2020_v3$ride_length ~ all_trips_2020_v3$member_casual + all_trips_2020_v3$day_of_week,
          FUN = mean)
# Notice that the days of the week are out of order. Let's fix that.
all_trips_2020_v3$day_of_week <- ordered(all_trips_2020_v3$day_of_week, levels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
# Find the average ride time by each day for members vs casual users
aggregate(all_trips_2020_v3$ride_length ~ all_trips_2020_v3$member_casual + all_trips_2020_v3$day_of_week,
          FUN = mean)
# analyze ridership data by type and day_of_week
all_trips_2020_v3 %>%
group_by(member_casual, day_of_week) %>% #groups by user type and day_of_week
  summarise(number_of_rides = n() #calculates the number of rides and average duration
            ,average_duration = mean(ride_length)) %>% # calculates the average duration
  arrange(member_casual, day_of_week) # sorts
# Visualize the number of rides by rider type
all_trips_2020_v3 %>%
  group_by(member_casual, day_of_week) %>%
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>%
  arrange(member_casual, day_of_week, na = TRUE) %>%
  ggplot(aes(x = day_of_week, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")
# Create a visualization for average duration
all_trips_2020_v3 %>%
  group_by(member_casual, day_of_week) %>%
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>%
  arrange(member_casual, day_of_week) %>%
  ggplot(aes(x = day_of_week, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge")
counts_2020 <- aggregate(all_trips_2020_v3$ride_length ~ all_trips_2020_v3$member_casual + all_trips_2020_v3$day_of_week, FUN = mean)

counts_2020_1 <- aggregate(all_trips_2020_v3$ride_length ~ all_trips_2020_v3$member_casual + all_trips_2020_v3$day_of_week + all_trips_2020_v3$rideable_type, FUN = mean)
  
#YEAR 2021
summary(all_trips_2021_v3$ride_length)
# Compare members and casual users
aggregate(all_trips_2021_v3$ride_length ~ all_trips_2021_v3$member_casual, FUN = mean)
aggregate(all_trips_2021_v3$ride_length ~ all_trips_2021_v3$member_casual, FUN = median)
aggregate(all_trips_2021_v3$ride_length ~ all_trips_2021_v3$member_casual, FUN = max)
aggregate(all_trips_2021_v3$ride_length ~ all_trips_2021_v3$member_casual, FUN = min)
# The average ride time by each day for members vs casual users
aggregate(all_trips_2021_v3$ride_length ~ all_trips_2021_v3$member_casual + all_trips_2021_v3$day_of_week,
          FUN = mean)
# Notice that the days of the week are out of order. Let's fix that.
all_trips_2021_v3$day_of_week <- ordered(all_trips_2021_v3$day_of_week, levels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
# Find the average ride time by each day for members vs casual users
aggregate(all_trips_2021_v3$ride_length ~ all_trips_2021_v3$member_casual + all_trips_2021_v3$day_of_week,
          FUN = mean)
# Analyze ridership data by type and day_of_week
all_trips_2021_v3 %>%
  group_by(member_casual, day_of_week) %>% #groups by user type and day_of_week
  summarise(number_of_rides = n() #calculates the number of rides and average duration
            ,average_duration = mean(ride_length)) %>% # calculates the average duration
  arrange(member_casual, day_of_week) # sorts
# Visualize the number of rides by rider type
all_trips_2021_v3 %>%
  group_by(member_casual, day_of_week) %>%
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>%
  arrange(member_casual, day_of_week, na = TRUE) %>%
  ggplot(aes(x = day_of_week, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")
# Create a visualization for average duration
all_trips_2021_v3 %>%
  group_by(member_casual, day_of_week) %>%
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>%
  arrange(member_casual, day_of_week) %>%
  ggplot(aes(x = day_of_week, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge")
counts_2021 <- aggregate(all_trips_2021_v3$ride_length ~ all_trips_2021_v3$member_casual + all_trips_2021_v3$day_of_week, FUN = mean)

counts_2021_1 <- aggregate(all_trips_2021_v3$ride_length ~ all_trips_2021_v3$member_casual + all_trips_2021_v3$day_of_week + all_trips_2021_v3$rideable_type, FUN = mean)


## EXPORT THE DATA
write.csv(counts_2019,"C://Users//User//Desktop//GOOGLE DATA ANALYTICS RESOURCES//.csv folder//avg_ride_length_2019.csv")
write.csv(counts_2020,"C://Users//User//Desktop//GOOGLE DATA ANALYTICS RESOURCES//.csv folder//avg_ride_length_2020.csv")
write.csv(counts_2021,"C://Users//User//Desktop//GOOGLE DATA ANALYTICS RESOURCES//.csv folder//avg_ride_length_2021.csv")
write.csv(counts_2020_1,"C://Users//User//Desktop//GOOGLE DATA ANALYTICS RESOURCES//.csv folder//avg_rideable_type_2020.csv")
write.csv(counts_2021_1,"C://Users//User//Desktop//GOOGLE DATA ANALYTICS RESOURCES//.csv folder//avg_rideable_type_2021.csv")


```


